---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# ewp

<!-- badges: start -->
[![R-CMD-check](https://github.com/pboesu/ewp/workflows/R-CMD-check/badge.svg)](https://github.com/pboesu/ewp/actions)
[![R-CMD-check](https://github.com/pboesu/ewp/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/pboesu/ewp/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

The goal of ewp is to provide a modelling interface for underdispersed count data based on the exponentially weighted Poisson (EWP) distribution described by Ridout & Besbeas (2004), allowing for nest-level covariates on the location parameter $\lambda$ of the EWP. Currently only the three-parameter version of the distribution (EWP_3) is implemented.

## Installation

You can install the development version of ewp directly from github like so:

``` r
remotes::install_github('pboesu/ewp')
```

## Example

The package contains a reconstructed version of the linnet dataset from Ridout & Besbeas (2004), which consists of 5414 clutch size records and is augmented with two synthetic covariates, one that is random noise, and one that is correlated with clutch size. The parameter estimates for the intercept-only model presented in Ridout & Besbeas (2004) can be reproduced like so:

```{r model1}
library(ewp)
fit_null <- ewp_reg(eggs ~ 1, data = linnet)
summary(fit_null)
```

Note that the linear predictor on $lambda$ uses a log-link.

A model with nest-level covariates can be fitted by specifying a more complex model formula - as in the base R `glm()`

```{r model2}
fit <- ewp_reg(eggs ~ cov1 + cov2, data = linnet)
summary(fit_null)
```

:warning: 
**Note that the maximum likelihood optimisation procedure is still experimental**

In particular: 

- Fitting is very slow (think minutes, not seconds), especially when estimating the Hessian matrix for more than a couple of parameters (think hours, not minutes)!
- Estimates may not be stable results for models with many covariates and/or very large sample sizes (1000s)

:warning:
