---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# ewp

<!-- badges: start -->
[![R-CMD-check](https://github.com/pboesu/ewp/workflows/R-CMD-check/badge.svg)](https://github.com/pboesu/ewp/actions)
[![Codecov test coverage](https://codecov.io/gh/pboesu/ewp/branch/main/graph/badge.svg)](https://app.codecov.io/gh/pboesu/ewp?branch=main)
<!-- badges: end -->

The goal of ewp is to provide a modelling interface for underdispersed count data based on the exponentially weighted Poisson (EWP) distribution described by Ridout & Besbeas (2004), allowing for nest-level covariates on the location parameter $\lambda$ of the EWP. Currently only the three-parameter version of the distribution (EWP_3) is implemented.

## Installation

You can install the development version of ewp directly from github like so (requires a working C++ compiler toolchain):

``` r
remotes::install_github('pboesu/ewp')
```

## Example

The package contains a reconstructed version of the linnet dataset from Ridout & Besbeas (2004), which consists of 5414 clutch size records and is augmented with two synthetic covariates, one that is random noise, and one that is correlated with clutch size. The parameter estimates for the intercept-only model presented in Ridout & Besbeas (2004) can be reproduced like so:

```{r model1}
library(ewp)
fit_null <- ewp_reg(eggs ~ 1, data = linnet)# this may take a few seconds
summary(fit_null)
```

Note that the linear predictor on $\lambda$ uses a log-link.

A model with nest-level covariates can be fitted by specifying a more complex model formula - as in the base R `glm()`

```{r model2}
fit <- ewp_reg(eggs ~ cov1 + cov2, data = linnet)# this may take 5-10 seconds
summary(fit)
```

Simulation based residual diagnostics are indirectly available through the package [DHARMa](https://cran.r-project.org/package=DHARMa), by using the `simulate.ewp` method:

```{r residuals}
library(DHARMa)
#simulate from fitted model
sims <- simulate(fit, nsim = 20)

#create a DHARMa abject
DH <- createDHARMa(simulatedResponse = as.matrix(sims),#simulated responses
                   observedResponse = linnet$eggs,#original response
                   fittedPredictedResponse = fit$fitted.values,#fitted values from ewp model
                   integerResponse = T)#tell DHARMa this is a discrete probability distribution

#plot diagnostics
plot(DH)
```

:warning: 
**Note that the maximum likelihood optimisation procedure is still experimental**
:warning:

In particular:    
- At the moment **the likelihood evaluation is optimised for small counts ($\lambda << 20) and uses a hard upper limit of 30 for potential counts**, this means the model is currently only suitable for datasets with expected counts up to 20-25, depending on the degree of underdispersion. A warning is issued if this criterion in not met when using `ewp_reg()`, but other functions may fail silently.
- Fitting is very slow (think minutes, not seconds), especially when estimating the Hessian matrix for more than a couple of parameters!
- Estimates may not be stable for models with many covariates and/or very large sample sizes (1000s). **Centering and scaling continuous covariates seems to help on that front.**

:warning::warning::warning:
